\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2014-07-17 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added the framework for the future option NUMBER_OF_SIMULTANEOUS_RUNS that I will add to setup/constants.h.in (off by default, i.e. equal to 1 by default) to have the ability to run several independent earthquakes in parallel from the same big job

@@

When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
specified in the main }\texttt{\small Par\_file}{\small , e.g. in
directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small . %As the files can be very large, there are several flags in the \texttt{\small Par\_file}
%that control the region in space and time that is saved. These are:
%\texttt{\small NTSTEP\_BETWEEN\_FRAMES}
The output is saved by each processor at the time interval specified
by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}

@@

We describe here in some detail a job submission procedure for the
Caltech 1024-node cluster, CITerra, under the LSF scheduling system.
We consider the submission of a regular forward simulation using the
internal mesher to create mesh partitions. The two main scripts are
\texttt{\small run\_lsf.bash}{\small , which compiles the Fortran
code and submits the job to the scheduler, and }\texttt{\small go\_mesher\_solver\_lsf\_basin.forward}{\small ,
which contains the instructions that make up the job itself. These
scripts can be found in }\texttt{\small utils/Cluster/lsf/}{\small{}
directory}{\small \par}

\section{Job submission \texttt{run\_lsf.bash}}

This script first sets the job queue to be `normal'. It then compiles
the mesher, database generator and solver together, figures out the
number of processors required for this simulation from the \texttt{DATA/Par\_file},
and submits the LSF job.
\begin{lyxcode}
\#!/bin/bash

\#~use~the~normal~queue~unless~otherwise~directed~queue=\textquotedbl{}-q~normal\textquotedbl{}~

if~{[}~\$\#~-eq~1~{]};~then

~~~~~~~~echo~\textquotedbl{}Setting~the~queue~to~\$1\textquotedbl{}

~~~~~~~~queue=\textquotedbl{}-q~\$1\textquotedbl{}~

fi~~\\
~~~\\
~\#~compile~the~mesher~and~the~solver~

d=`date'~echo~\textquotedbl{}Starting~compilation~\$d\textquotedbl{}~

make~clean~

make~xmeshfem3D~

make~xgenerate\_databases~

make~xspecfem3D~

d=`date'~

echo~\textquotedbl{}Finished~compilation~\$d\textquotedbl{}~~\\
~~~\\
~\#~get~total~number~of~nodes~needed~for~solver~

NPROC=`grep~NPROC~DATA/Par\_file~|~cut~-c~34-~'~~\\
~~~\\

\#~compute~total~number~of~nodes~needed~for~mesher~

NPROC\_XI=`grep~NPROC\_XI~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\

NPROC\_ETA=`grep~NPROC\_ETA~DATA/meshfem3D\_files/Mesh\_Par\_file~|~cut~-c~34-~'~~~~\\
~\#~total~number~of~nodes~is~the~product~of~the~values~read~

numnodes=\$((~\$NPROC\_XI~{*}~\$NPROC\_ETA~))~~\\
~~~\\

\#~checks~total~number~of~nodes~

if~{[}~\$numnodes~-neq~\$NPROC~{]};~then

~~~~~~~~echo~\textquotedbl{}error~number~of~procs~mismatch\textquotedbl{}

~~~~~~~~exit~

fi~~\\

~~\\
~echo~\textquotedbl{}Submitting~job\textquotedbl{}~

bsub~\$queue~-n~\$numnodes~-W~60~-K~<go\_mesher\_solver\_lsf.forward~
\end{lyxcode}

\section{Job script \texttt{go\_mesher\_solver\_lsf.forward}}

This script describes the job itself, including setup steps that can
only be done once the scheduler has assigned a job-ID and a set of
compute nodes to the job, the \texttt{run\_lsf.bash} commands used
to run the mesher, database generator and the solver, and calls to
scripts that collect the output seismograms from the compute nodes
and perform clean-up operations.
\begin{enumerate}
\item First the script directs the scheduler to save its own output and
output from \texttt{stdout} into ~\\
 \texttt{\small OUTPUT\_FILES/\%J.o}{\small , where }\texttt{\small \%J}{\small{}
is short-hand for the job-ID; it also tells the scheduler what version
of }\texttt{\small mpich}{\small{} to use (}\texttt{\small mpich\_gm}{\small )
and how to name this job (}\texttt{\small go\_mesher\_solver\_lsf}{\small ). }{\small \par}
\item The script then creates a list of the nodes allocated to this job
by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
and parsing the output into a one-column list using the Perl script
\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates
a set of scratch directories on these nodes (\texttt{\small /scratch/}{\small ~}\\
{\small{} }\texttt{\small \$USER/DATABASES\_MPI}{\small ) to be used
as the }\texttt{\small LOCAL\_PATH}{\small{} for temporary storage
of the database files. The scratch directories are created using }\texttt{\small shmux}{\small ,
a shell multiplexor that can execute the same commands on many hosts
in parallel. }\texttt{\small shmux}{\small{} is available from Shmux
\urlwithparentheses{web.taranis.org/shmux/}. Make sure that the
}\texttt{\small LOCAL\_PATH}{\small{} parameter in }\texttt{\small DATA/Par\_file}{\small{}
is also set properly. }{\small \par}
\item The next portion of the script launches the mesher, database generator
and then the solver using \texttt{run\_lsf.bash}.
\item The final portion of the script collects the seismograms and performs
clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
and \texttt{cleanmulti.pl}. \end{enumerate}
\begin{lyxcode}
\#!/bin/bash~-v~

\#BSUB~-o~OUTPUT\_FILES/\%J.o~

\#BSUB~-a~mpich\_gm~

\#BSUB~-J~go\_mesher\_solver\_lsf~~\\
~~~\\
~\#~set~up~local~scratch~directories

BASEMPIDIR=/scratch/\$USER/DATABASES\_MPI

mkdir~-p~OUTPUT\_FILES~

echo~\textquotedbl{}\$LSB\_MCPU\_HOSTS\textquotedbl{}~>~OUTPUT\_FILES/lsf\_machines~

echo~\textquotedbl{}\$LSB\_JOBID\textquotedbl{}~>~OUTPUT\_FILES/jobid

remap\_lsf\_machines.pl~OUTPUT\_FILES/lsf\_machines~>~OUTPUT\_FILES/machines

shmux~-M50~-Sall~-c~\textquotedbl{}rm~-r~-f~/scratch/\$USER;~\textbackslash{}

~~~~~~mkdir~-p~/scratch/\$USER;~mkdir~-p~\$BASEMPIDIR\textquotedbl{}~\textbackslash{}

~~~~~~-~<~OUTPUT\_FILES/machines~>/dev/null~~\\
~~~\\
~\#~run~the~specfem~program

current\_pwd=\$PWD~~\\

cd~bin/~~\\

run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xmeshfem3D~~\\

run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xgenerate\_databases~~\\

run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~\$current\_pwd/xspecfem3D~~\\
~~~\\
~\#~collect~seismograms~and~clean~up

cd~current\_pwd/~~~\\

mkdir~-p~SEM

cd~SEM/~

collect\_seismo.pl~../OUTPUT\_FILES/lsf\_machines

cleanbase.pl~../OUTPUT\_FILES/machines
\end{lyxcode}

@@

%% magnoni 6/12

>>>>

We have also added the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of few to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.
To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS
to a value greater than 1 in file setup/constants.h.in before
configuring and compiling the code.
When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
all the individual runs must use the same number of processor cores
(which as usual is NPROC in the input file DATA/Par\_file,
and thus the total number of processor cores to request from the batch system
should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits).

@@

When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
specified in the main }\texttt{\small Par\_file}{\small , e.g. in
directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
The output is saved by each processor at the time interval specified
by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}

@@

@@
